Taking Linear Logic Apart

### 2018-04-28 doc/tlla2018/main.tex
Added hyper-environment to (T) rule; Minor fixes

\documentclass[draft,submission,copyright,creativecommons]{eptcs}
\providecommand{\event}{Linearity and TLLA 2018}
\usepackage[british]{babel}
\usepackage[all]{foreign}
\usepackage[fleqn]{amsmath}
\usepackage{amsthm}
\allowdisplaybreaks
\usepackage[inline]{enumitem}
\usepackage{breakurl} % Not needed if you use pdflatex only.
\usepackage{underscore} % Only needed if you use pdflatex.
\usepackage{textcomp,textgreek,upgreek}
\usepackage{nameref,hyperref}
\usepackage[nameinlink,noabbrev]{cleveref}
\hypersetup{hidelinks,final}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{example}{Example}
\input{preamble-typing}
\usepackage[sectionbib,square,numbers,sort&compress]{natbib}
\renewcommand*{\eg}{e.g.\@\xspace}
\renewcommand*{\ie}{i.e.\@\xspace}
\usepackage{more-space}
\usepackage{setspace}
\renewcommand{\baselinestretch}{0.925}

@@

\paragraph{Types}
We use the same definitions for types and environments for \hcp as we used for \cp.
However, we introduce a new layer on top of sequents: hypersequents.
As \cp is a one-sided logic, and it uses the left-hand side of the
turnstile to write the process, the traditional hypersequent notation can look
confusing: ``$\seq[P]{\Gamma_1}\hsep\dots\hsep\seq{\Gamma_n}$'' seems to claim
that $\tm{P}$ acts according to protocol $\ty{\Gamma_1}$.
What are all the other $\Gamma$s doing there? Are they typing empty processes?
Therefore, we opt to leave out the repeated turnstile, and instead work with the
notion of ``hyper-environments''. However, we will still refer to our system as a
hypersequent system.
Hyper-environments are either empty, or consist of a series of typing
environments, separated by vertical bars.
A hyper-environment $\ty{\Gamma_1 \hsep \dots \hsep \Gamma_n}$ types a series of
$n$ entangled, but independent processes.
\begin{definition}[Hyper-environment]\label{def:hcp-hyper-environment}
  \(
  \ty{\mathcal{G}}, \ty{\mathcal{H}} :=
  \ty{\emptyhypercontext}
  \; \mid \;
  \ty{\Gamma_1 \hsep \dots \hsep \Gamma_n}
  \)
\end{definition}\noindent
Typing judgements in \hcp associate processes with hyper-environments.
\textsc{H-Mix} composes two processes in parallel, but remembers that they are
independent on in the sequent.
\textsc{H-Cut} and $(\tens)$ take as their premise a process which consists of
at least two unconnected processes, and connects them, eliminating the vertical
bar.
Note that each logical rule has the side condition that $\tm{x} \not\in
\mathcal{G}$, which can be read as ``you cannot act on one end-point of $x$ if
you are also holding its other end-point''.
This prevents self-locking processes, \eg
$\tm{\piBoundSend{x}{}{\piRecv{x}{}{\piHalt}}}$.
\begin{definition}[Typing judgements]\label{def:hcp}
  A typing judgement $\seq[P]{\Gamma_1 \hsep \dots \hsep \Gamma_n}$ denotes
  that the process $\tm{P}$ consists of $n$ independent, but potentially
  entangled processes, each of which communicates according to its own
  protocol $\Gamma_i$.
  Typing judgements can be constructed using the inference rules in \cref{fig:hcp}.
  Furthermore, each logical rule has the side condition that $\tm{x} \not\in
  \ty{\mathcal{G}}$.
\end{definition}\noindent
\begin{figure}[htb]
  \begin{center}\hccpInfAx \hccpInfCycle\end{center}
  \vspace*{-0.5\baselineskip}
  \begin{center}\hccpInfMix \hccpInfHalt\end{center}
  \vspace*{-0.5\baselineskip}
  \begin{center}\hccpInfBoundTens \hccpInfParr \hccpInfOne\end{center}
  \vspace*{-0.5\baselineskip}
  \begin{center}\hccpInfBot \hccpInfPlus1 \hccpInfPlus2\end{center}
  \vspace*{0.5\baselineskip}
  \begin{center}\hccpInfWith \hccpInfNil \hccpInfTop\end{center}
  \caption{Hypersequent Classical Processes, typing.}
  \label{fig:hcp}
\end{figure}

@@

\looseness=-1
The idea of using hypersequents to capture parallelism in linear logic judgements is not
novel: \citet{carbone2018} extended the multiplicative-additive fragment
of intuitionistic linear logic with hypersequents to type global descriptions of
process communications known as choreographies.
This work is distinct from our approach in that \hcp is based on classical linear
logic and manipulates hypersequents differently: in \citet{carbone2018}, hypersequents can be formed only when sequents share
resources (\cf \textsc{H-Mix}), and resource sharing is then explicitly tracked
using an additional connection modality (which is not present in \hcp).

%\section{Conclusions}
%\label{sec:conclusions}

>>>>

\documentclass[draft,submission,copyright,creativecommons]{eptcs}
\providecommand{\event}{Linearity and TLLA 2018}
\usepackage[british]{babel}
\usepackage[all]{foreign}
\usepackage[fleqn]{amsmath}
\usepackage{amsthm}
\allowdisplaybreaks
\usepackage[inline]{enumitem}
\usepackage{breakurl} % Not needed if you use pdflatex only.
\usepackage{underscore} % Only needed if you use pdflatex.
\usepackage{textcomp,textgreek,upgreek}
\usepackage{nameref,hyperref}
\usepackage[nameinlink,noabbrev]{cleveref}
\hypersetup{hidelinks,final}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{example}{Example}
\input{preamble-typing}
\usepackage[sectionbib,square,numbers,sort&compress]{natbib}
\renewcommand*{\eg}{e.g.\@\xspace}
\renewcommand*{\ie}{i.e.\@\xspace}
\usepackage{more-space}
\usepackage{setspace}
\everypar{\looseness=-1}
\renewcommand{\baselinestretch}{0.975}

@@

\paragraph{Types}
We use the same definitions for types and environments for \hcp as we used for \cp.
However, we introduce a new layer on top of sequents: hypersequents.
As \cp is a one-sided logic, and it uses the left-hand side of the
turnstile to write the process, the traditional hypersequent notation can look
confusing: ``$\seq[P]{\Gamma_1}\hsep\dots\hsep\seq{\Gamma_n}$'' seems to claim
that $\tm{P}$ acts according to protocol $\ty{\Gamma_1}$.
What are all the other $\Gamma$s doing there? Are they typing empty processes?
Therefore, we opt to leave out the repeated turnstile, and instead work with the
notion of ``hyper-environments''. However, we will still refer to our system as a
hypersequent system.
Hyper-environments are either empty, or consist of a series of typing
environments, separated by vertical bars.
A hyper-environment $\ty{\Gamma_1 \hsep \dots \hsep \Gamma_n}$ types a series of
$n$ entangled, but independent processes.
\begin{definition}[Hyper-environment]\label{def:hcp-hyper-environment}
  \(
  \ty{\mathcal{G}}, \ty{\mathcal{H}} :=
  \ty{\emptyhypercontext}
  \; \mid \;
  \ty{\Gamma_1 \hsep \dots \hsep \Gamma_n}
  \)
\end{definition}\noindent
Typing judgements in \hcp associate processes with hyper-environments.
\textsc{H-Mix} composes two processes in parallel, but remembers that they are
independent in the sequent.
\textsc{H-Cut} and $(\tens)$ take as their premise a process which consists of
at least two independent processes, and connects them, eliminating the vertical
bar.
Note that each logical rule has the side condition that $\tm{x} \not\in
\ty{\mathcal{G}}$, which can be read as ``you cannot act on one end-point of $x$
if you are also holding its other end-point''.
This prevents self-locking processes, \eg
$\tm{\piBoundSend{x}{}{\piRecv{x}{}{\piHalt}}}$.
\begin{definition}[Typing judgements]\label{def:hcp}
  A typing judgement $\seq[P]{\Gamma_1 \hsep \dots \hsep \Gamma_n}$ denotes
  that the process $\tm{P}$ consists of $n$ independent, but potentially
  entangled processes, each of which communicates according to its own
  protocol $\Gamma_i$.
  Typing judgements can be constructed using the inference rules in \cref{fig:hcp}.
  Furthermore, each logical rule has the side condition that $\tm{x} \not\in
  \ty{\mathcal{G}}$.
\end{definition}\noindent
\begin{figure}[htb]
  \begin{center}\hccpInfAx \hccpInfCycle\end{center}
  \vspace*{-0.5\baselineskip}
  \begin{center}\hccpInfMix \hccpInfHalt\end{center}
  \vspace*{-0.5\baselineskip}
  \begin{center}\hccpInfBoundTens \hccpInfParr \hccpInfOne\end{center}
  \vspace*{-0.5\baselineskip}
  \begin{center}\hccpInfBot \hccpInfPlus1 \hccpInfPlus2\end{center}
  \vspace*{0.5\baselineskip}
  \begin{center}\hccpInfWith \hccpInfNil \hccpInfTop\end{center}
  \caption{Hypersequent Classical Processes, typing.}
  \label{fig:hcp}
\end{figure}

@@

The idea of using hypersequents to capture parallelism in linear logic
judgements is not novel: \citet{carbone2018} extended the
multiplicative-additive fragment of intuitionistic linear logic with
hypersequents to type global descriptions of process communications known as
choreographies.
This work is distinct from our approach in that \hcp is based on classical
linear logic and manipulates hypersequents differently: in \citet{carbone2018},
hypersequents can be formed only when sequents share resources (\cf
\textsc{H-Mix}), and resource sharing is then explicitly tracked using an
additional connection modality (which is not present in \hcp).

\section{Conclusions}
\label{sec:conclusions}
